Making Sequential Programs Deterministic

### 2019-03-28 design.tex
Edits in sections 2.3 and 2.4

@@

    $\infer[\rulename{nondet collection}]{\wellformed{\|NonDet| \ \beta_c\angles{\|NonDet| \ \beta_e}}}{\beta_c \sqsubseteq \<Collection>}$

@@

    $\infer[\rulename{nondet map}]{\wellformed{\|NonDet| \ \beta_m\angles{\|NonDet| \ \beta_k, \|NonDet| \ \beta_v}}}{\beta_m \sqsubseteq \<Map>}$

    \caption{Type well formedness}
\todo{The \rulename{noncollection} rule is not correct, because not every type is
  well-formed.  It needs an antecedent stating that $\beta$ is not a
  collection or map.}

@@

\OurTypeSystem supports three types of polymorphism:  type
polymorphism, basetype polymorphism, and qualifier polymorphism.
These apply to both classes and methods.
\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}

>>>>

\todo{Define the Greek.  For example:

  A type $\tau = \kappa\ \beta$ consists of a qualifier and a basetype.
  Neither may be omitted.

  $\beta$ is a type that lacks its type qualifier, but has type qualifiers
  on any type arguments:
  $\beta = C | C\angles{\overline{\tau}}$.
  }

@@

    $\infer[\rulename{nondet collection}]
    {\wellformed{\|NonDet| \ C\angles{\|NonDet| \ \beta_e}}}
    {C \sqsubseteq \<Collection> & \wellformed{\|NonDet| \ \beta_e}}$

@@

    $\infer[\rulename{nondet map}]
    {\wellformed{\|NonDet| \ \beta_m\angles{\|NonDet| \ \beta_k, \|NonDet| \ \beta_v}}}
    {\beta_m \sqsubseteq \<Map> & \wellformed{\|NonDet| \ \beta_k} & \wellformed{\|NonDet| \ \beta_v}}$

    \caption{Type well formedness}

@@

\OurTypeSystem supports three types of polymorphism:  type
polymorphism, basetype polymorphism, and qualifier polymorphism.
These apply to both classes and methods.
\todo{Greekify the following}
\begin{itemize}
\item
In \theDeterminismCheckerImplementation,
type polymorphism is handled by Java's generics mechanism, which
\theDeterminismChecker fully supports, including class and method generics,
inference, etc.
Given the Java declaration
\begin{Verbatim}
<T> T identity(T arg) { return arg; }
\end{Verbatim}
the type of \<identity(arg)> is the same as the type of
\<arg>, and a programmer can write a type such as \codeid{List<@NonDet
  Integer>} (but see below for details about collections).
\item
Basetype polymorphism is enabled by writing a type qualifier on a use of a
type variable, which overrides the type qualifier at the instantiation
site.
For example, a \<choose> operation on sets could be defined in Java as
\begin{Verbatim}
class ChooseableSet<T> implements Set<T> {
  // return an arbitrary element from the set
  @NonDet T choose();
}
\end{Verbatim}
\item
Java does not provide a syntax that can be used for qualifier polymorphism,
so we use a special type qualifier name, \<@PolyDet>.
A qualifier-polymorphic method \<m> with signature $\forall \kappa . \kappa\ \<int> \times \<Det boolean> \rightarrow
\kappa\ \<String>$ would be written in our dialect of Java as
\begin{Verbatim}
@PolyDet String m(@PolyDet int, @Det boolean)
\end{Verbatim}
Each use of \<@PolyDet> stands for a use of the qualifier variable
$\kappa$, and there is no need to declare the qualifier variable $\kappa$.
\TheDeterminismChecker currently supports
qualifier
polymorphism on methods but not on classes.
Therefore, \<@PolyDet> may be written in methods (signatures and bodies)
but not on fields.
\end{itemize}
